---
title: "Genome List"
author: '9290440'
date: "2 August 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library("readxl")
library("tidyverse")
library("kableExtra")
library(forcats)
library(gapminder)
library(writexl)
library(RColorBrewer)
```

Call in final genome list

```{r}
genomes <- read_excel("Final_Genome_Set.xlsx")
#And now to see the distribution of the genomes I have collected 
phylum.graph.new <- ggplot(genomes, mapping = aes(fct_infreq(factor(Phylum)))) + geom_bar() + coord_flip() + ggtitle("Count of Isolate Genomes for each Phylum") + xlab("Phylum") + ylab("Count of Isolates") + geom_text(stat = 'count',aes(label =..count.., hjust = 0.1))
phylum.graph.new

genomes <- mutate_at(genomes, vars( "Phylum" ,"Class", "Order", "Family", "Genus", "Species"), as.factor)
```

Now to do some investigation as to how I can further filter genomes greater than 51 isolates. I want to make the large phyla like Proteobacteria a lot smaller so that my results will be less biased towards the larger phyla. To do that I will use PIPES!! to create tables of the individual phyla in question and make some summary table of the distribution of class and order within each phyla

Some tables displaying Proteobacteria distribution in class and then order. Going by the rough guideline above of having no more than ~50 isolates per phyla, one isolate per Order fits quite well here.

One isolate per Order will yield 51 isolates

```{r}
genomes %>% filter(Phylum == "Proteobacteria") %>% group_by(Class) %>% tally()
genomes %>% filter(Phylum == "Proteobacteria") %>% group_by(Order) %>% tally() 
genomes.proteo <- genomes %>% filter(Phylum == "Proteobacteria") %>% filter(`Scaffold Count   * assembled` == 1)
genomes.proteo %>% group_by(Order) %>% tally()
genomes.proteo.filt <- genomes.proteo %>% group_by(Order) %>% sample_n(1)
```

Firmicutes distribution

One isolate per family will yeild 46 isolates

```{r}
genomes %>% filter(Phylum == "Firmicutes") %>% group_by(Class) %>% tally()
genomes %>% filter(Phylum == "Firmicutes") %>% group_by(Order) %>% tally()
genomes %>% filter(Phylum == "Firmicutes") %>% group_by(Family) %>% tally()

genomes.firm <- genomes %>% filter(Phylum == "Firmicutes") %>% filter(`Scaffold Count   * assembled` == 1)
genomes.firm %>% group_by(Family) %>% tally()
genomes.firm.filt <- genomes.firm %>% group_by(Family) %>% sample_n(1)
```

Actinobacteria distribution

One isolate per Family will yield 50 isolates

```{r}
genomes %>% filter(Phylum == "Actinobacteria") %>% group_by(Class) %>% tally()
genomes %>% filter(Phylum == "Actinobacteria") %>% group_by(Order) %>% tally()
genomes %>% filter(Phylum == "Actinobacteria") %>% group_by(Family) %>% tally()

genomes.act <- genomes %>% filter(Phylum == "Actinobacteria") %>% filter(`Scaffold Count   * assembled` == 1)
genomes.act %>% group_by(Family) %>% tally()
genomes.act.filt <- genomes.act %>% group_by(Family) %>% sample_n(1)
```

Bacteroidetes distribution

Ok so Family contains 31 families and Genus has 106 genera

Two isolate per Family will yield ~60 isolates

```{r}
genomes %>% filter(Phylum == "Bacteroidetes") %>% group_by(Class) %>% tally()
genomes %>% filter(Phylum == "Bacteroidetes") %>% group_by(Order) %>% tally()
genomes %>% filter(Phylum == "Bacteroidetes") %>% group_by(Family) %>% tally()
genomes %>% filter(Phylum == "Bacteroidetes") %>% group_by(Genus) %>% tally()

genomes.bact <- genomes %>% filter(Phylum == "Bacteroidetes") %>% filter(`Scaffold Count   * assembled` == 1)
genomes.bact %>% group_by(Family) %>% tally()
genomes.bact.filt <- genomes.bact %>% group_by(Family) %>% sample_n(1)
```

Tenericutes distribution

Alright, this one is a bit more difficult to cut back on. Genus results in 8 groups and the Species results in 86 groups. So I think the best way around this would be to randomly remove 5 from the Spiroplasma genus and 25 from the Mycoplasma genus so the distribution is a bit more even and will still result in ~50 isolate (will result in 54)

```{r}
genomes %>% filter(Phylum == "Tenericutes") %>% group_by(Class) %>% tally()
genomes %>% filter(Phylum == "Tenericutes") %>% group_by(Order) %>% tally()
genomes %>% filter(Phylum == "Tenericutes") %>% group_by(Family) %>% tally()
genomes %>% filter(Phylum == "Tenericutes") %>% group_by(Genus) %>% tally()

genomes.ten <- genomes %>% filter(Phylum == "Tenericutes") %>% filter(`Scaffold Count   * assembled` == 1)
genomes.ten %>% group_by(Family) %>% tally()
genomes.ten.1 <- genomes.ten %>% subset(genomes.ten$Family == "unclassified")
genomes.ten <- genomes.ten[-c(74),]
genomes.ten.filt <- genomes.ten %>% group_by(Family) %>% sample_n(10)
```

Cyanobacteria distribution

One isolate per Genus will yield 51 isolates

```{r}
genomes %>% filter(Phylum == "Cyanobacteria") %>% group_by(Class) %>% tally()
genomes %>% filter(Phylum == "Cyanobacteria") %>% group_by(Order) %>% tally()
genomes %>% filter(Phylum == "Cyanobacteria") %>% group_by(Family) %>% tally()
genomes %>% filter(Phylum == "Cyanobacteria") %>% group_by(Genus) %>% tally()

genomes.cyan <- genomes %>% filter(Phylum == "Cyanobacteria")
genomes.cyan %>% group_by(Family) %>% tally()
genomes.cyan.filt <- genomes.cyan %>% group_by(Family) %>% sample_n(1)
```

OK!! After some more painful flitering, we have the final genome collection!

```{r}
Complete.genomes = read_excel("Complete genome.xlsx")

Complete.genomes <- mutate_at(Complete.genomes, vars( "Phylum" ,"Class", "Order", "Family", "Genus", "Species"), as.factor)

phylum.graph.final <- ggplot(Complete.genomes, mapping = aes(fct_infreq(factor(Phylum)))) + 
  geom_bar() + 
  xlab("Phylum") + 
  ylab("Count of Isolates") + 
  geom_text(stat = 'count',aes(label =..count.., hjust = 0.5, vjust = -0.5)) + 
  theme(axis.text.x = element_text(angle = 60, hjust = 1.1, vjust = 1)) +
  ylim (0, 60) + 
  theme_classic()
phylum.graph.final + theme(axis.text.x = element_text(angle = 60, hjust = 1.1, vjust = 1))
ggsave("Plot of genome isolates per Phylum", plot = last_plot(), device = "pdf", width = 8, height = 11)

phylum.graph.poster <- ggplot(Complete.genomes, mapping = aes(fct_infreq(factor(Phylum)))) + geom_bar() + xlab("Phylum") + ylab("Count of Isolates") + geom_text(stat = 'count',aes(label =..count.., hjust = -0.5, vjust = 0.5, size = 40))  +ylim (0, 60) + theme_classic()

abc <- phylum.graph.poster + 
  coord_flip()  + 
  theme(axis.text.y = element_text(size = 16),
        axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 20))

ggsave(
  "complete.isolate.poster.pdf",
  plot = last_plot(),
  device = "pdf",
  path = NULL,
  scale = 1,
  width = 15,
  height = 10,
  units = "in",
  limitsize = TRUE
)



ggsave("Plot of genome isolates per Phylum - poster.jpeg", plot = last_plot(), device = "jpeg", width = 3.8, height = 2.4)

unique(Complete.genomes$Phylum)

```


```{r}
library(readxl)
library(tidyverse)
library(RColorBrewer)
poster <- read_xlsx("Book1.xlsx")

poster$Phylum <- as.factor(as.character(poster$Phylum))
poster$`Isolate count` <- as.numeric(as.character(poster$`Isolate count`))
class(poster$`Isolate count`)

ggplot(poster, aes(x="", y=Phylum, fill=`Isolate count`)) +
  geom_bar(stat="identity", width=1) +
  coord_polar("y", start=0)

ggplot(poster, aes(x="", y=Phylum, fill=`Isolate count`)) +
  geom_bar(stat="identity", width=1, color="white") +
  coord_polar("y", start=0) +
  
  theme_void() # remove background, grid, numeric labels
myPalette <- brewer.pal(6, "Set2")

pie(poster$`Isolate count`, labels = c("Actinobacteria", "Bacteroidetes", "Cyanobacteria", "Firmicutes", "Proteobacteria", "Other 150 phylum"), col = myPalette, border = "white", clockwise = T )
```




